Time Series Modeling of Varying Complexity Captures Subdiffusive Solute Dynamics and Predicts Long Timescale Behavior in Nanoscale Pores.

### 2019-12-12 Ben_Manuscripts/stochastic_transport/Draft.tex
Added comments on flux sections.

  We determined the rate at which solutes cross macroscopic-length pores based on the
  Hill relation:~\cite{hill_free_1989}
  \begin{equation}
  J = \frac{1}{MFPT}
  \label{eqn:hill_relation}
  \end{equation}
  where $J$ is the single particle solute flux and MFPT refers to the mean first passage
  time. For particle concentrations higher than 1 at the pore entrance, one can multiply
  Equation~\ref{eqn:hill_relation} by the total number of particles to get the total flux.
  In the context of our work, the MFPT describes the average length of time it takes a particle
  to move from the pore entrance to the pore exit.

@@

  To come up with an analytical equation describing the passage time distributions, 
  one can frame the problem as a pulse of particles instantaneously released at the 
  pore inlet ($z=0$) which moves at a constant velocity and spreads out as it 
  approaches $L$. This gives results equivalent to if we had released each particle
  individually and then analyzed the positions of the ensemble of trajectories as 
  a function of time.

  To derive an analytical equation, first consider an initial pulse spreading out 
  over time with a fixed mean. We can solve for the time-dependent probability density
  of particle positions by solving the one dimensional diffusion equation:
  \begin{equation}
  \frac{\partial p}{\partial t} = D \frac{\partial^2 p}{\partial z^2}
  \end{equation}
  The appropriate initial and boundary conditions are:
  $$BC1: t > 0, z = \infty, p = 0$$
  $$BC2: t > 0, z = 0, \frac{\partial p}{\partial z} = 0$$
  $$IC: t = 0, c = M\delta(z)$$
  where $M$ is the total number of particles in the system. It has been shown 
  elsewhere that the solution to this equation is:~\cite{cussler_diffusion:_2009}
  \begin{equation}
  p(z, t) = \frac{1}{\sqrt{4 \pi D t}}exp\bigg(\frac{-z^2}{4Dt}\bigg)
  \end{equation}
  We can make the substitution $z = z - vt$, where $v$ represents a constant average
  velocity, in order to linearly shift the mean as a function of time:
  \begin{equation}
  p(z, t) = \frac{1}{\sqrt{4 \pi D t}}exp\bigg(\frac{-(z - vt)^2}{4Dt}\bigg)
  \end{equation}
  One can track the fraction of particles, $F$, that have crossed the pore 
  boundary by integrating:
  \begin{equation}
  F(t) = \int_L^\infty p~dz = erfc\bigg(\frac{L - vt}{2\sqrt{D t}}\bigg)
  \end{equation}
  where $L$ is the pore length. This represents the cumulative first passage 
  time distribution so we take its derivative in order to arrive at the first
  passage time distribution:
  \begin{equation}
  P(t) = -\frac{1}{\sqrt{\pi}}e^{-(L - vt)^2 / (4Dt)}\bigg(-\frac{D(L - vt)}{4(Dt)^{3/2}} - \frac{v}{2\sqrt{Dt}}\bigg)
  \label{eqn:passage_times}
  \end{equation} 
  where the only free parameters for fitting are $v$ and $D$. We calculated the
  expected value of Equation~\ref{eqn:passage_times} in order to get the MFPT.

@@

  %BJC4: not sure if the next couple paragraphs belong in methods
  It is computationally unfeasible to simulate trajectories long enough that they
  traverse the length of a macroscopic pore. To date, the thinnest H\textsubscript{II}
  LLC membrane synthesized with the monomer in this work was 7$\mu$m thick. Using
  24 cores to simulate trajectory realizations in parallel, it takes on the order 
  of 1 day to simulate 50000 realizations of solutes traversing a 50 nm pore. The 
  RAM requirements and performance scales greater than linearly and thus would 
  take an unfeasible amount of memory and time to simulate transport through a 
  pore over 100 times longer.

  %BJC4: methods?
  We used simulated trajectories which traverse computationally-reasonable length
  pores in order to construct an empirical model which one can use to estimate 
  particle flux for arbitrary length pores. As shown in Figure~\ref{fig:flux_curves},
  the flux appears to scale according to a power law of the form:
  \begin{equation}
  J(L) = cL^{-\beta} 
  \label{eqn:flux_decay}
  \end{equation}
  %while the MFPT scales as its inverse, $(1/c)L^{\beta}$.

  The prefactor, $c$, in Equation~\ref{eqn:flux_decay} grows with increasing
  solute MSD while the value of the growth exponent, $\beta$, is inversely 
  related to the Hurst parameter, $H$ (see Figure~\ref{fig:flux_parameters}).
  In Figure~\ref{fig:flux_curves_brownian}, we removed the correlation 
  structure (set $H$=0.5) from AD trajectory realizations so that all hops 
  were uncorrelated. Regardless of the parameters of the dwell time distribution,
  the growth exponent is always about 2, the same value that we observe for 
  Brownian motion (see Figure~\ref{S-fig:mfpt_curve_brownian} of the 
  Supporting Information).

@@

@@

  %BJC: need simulations with increased length for L=35-50 nm pores of acetic acid and methanol
  %BJC: see of clauset has better way of fitting t^-alpha (alpha might be too big)
  %BJC: not sure mfpt is necessary. Takes up space.
  %BJC: change alpha to beta
  %BJC: make a chart with parameters
  %BJC: should have a paragraph about c -- roughly follows their MSDs.
  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{flux_curves.pdf}
  \caption{Fits to $cL^{-\beta}$}\label{fig:flux_curves_regular}
  \end{subfigure}
  % BJC: could use better color scheme. Maybe outline hurst parameters
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{flux_parameters.pdf}
  \caption{}\label{fig:flux_parameters}
  \end{subfigure}
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{mfpt_curves.pdf}
%  \caption{}\label{fig:mfpt_curves}
%  \end{subfigure}
  %BJC: going out to 3 decimal places seems silly
  %BJC: might get closer to beta=2 if I use different curve fitting technique than fitting line to log-log
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{flux_curves_brownian.pdf}
  \caption{Fixed $H=0.5$}\label{fig:flux_curves_brownian}
  \end{subfigure}
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{mfpt_curves_brownian.pdf}
%  \caption{}\label{fig:mfpt_curves_brownian}
%  \end{subfigure}
  \caption{(a) The single particle solute flux decays with increasing pore length. 
  Solutes with high MSDs have higher fluxes. (b) We fit the single particle solute flux 
  versus pore length, $L$, to a power law function of the form $cL^{-\beta}$ (dashed 
  lines in (a)). Solutes with higher fluxes have higher values of $c$. Solutes with 
  smaller Hurst parameters, $H$, have larger values of $\beta$. (c) If we remove 
  correlation between hops (set $H$=0.5), the flux increases significantly and 
  $\beta$ is constant at about 2 for all solutes independent of dwell distribution
  parameters.}\label{fig:flux_curves}
  \end{figure}

@@

  In Figure~\ref{fig:selectivity}, we plot Equation~\ref{eqn:selectivity_ratio} for pore
  lengths ranging from those studied in Figure~\ref{fig:flux_curves} to macroscopic-length
  pores. Greater differences in $\beta$ values leads to selectivities that are stronger 
  functions of pore length. For $(c_i / c_j)=1$, LLC membranes will be more selective
  towards passage of solutes with lower $\beta$ values versus those with high $\beta$ values.
  When $\beta$ values are equal, as is the case for uncorrelated motion, selectivity only
  depends on the pre-factors, $c$.

@@

  Out of the solutes studied, the data suggests that this particular LLC membrane 
  would be most useful for selectively separating ethylene glycol from methanol and 
  acetic acid. 
  %It also shows reasonable selectivity of Urea over acetic acid.  
  Ethylene glycol has the highest $c$ value and the lowest $\beta$ value. The high
  value of $c$ itself greatly enhances its selective passage. Even more importantly,
  the relatively large difference between its $\beta$ value and those of methanol 
  and acetic acid give rise to a strong pore length dependence. The selectivity is
  enhanced as the pore length increases at the cost of lower solute flux, a 
  phenomenon ubiquitous in polymer separations membranes.~\cite{geise_water_2011}

@@

    Neglects concentration polarization, membrane defects and fouling.

>>>>

  We determined the rate at which solutes cross macroscopic-length pores based on the
  Hill relation:~\cite{hill_free_1989}
  \begin{equation}
  J = \frac{1}{MFPT}
  \label{eqn:hill_relation}
  \end{equation}
  where $J$ is the single particle solute flux and MFPT refers to the mean first passage
  time. 
  %MRS5: I would phrase below: 
  To account for input concentration dependence of the flux, assuming that 
  particles are uncorrelated, 
  %For particle concentrations higher than 1 at the pore entrance, 
  one can multiply
  Equation~\ref{eqn:hill_relation} by the total number of particles to get the total flux.
  In the context of our work, the MFPT describes the average length of time it takes a particle
  to move from the pore entrance to the pore exit.

@@

  To come up with an analytical equation describing the passage time distributions, 
  one can frame the problem as a pulse of particles instantaneously released at the 
  pore inlet ($z=0$) which moves at a constant velocity and spreads out as it 
  approaches $L$. This gives results equivalent to if we had released each particle
  individually and then analyzed the positions of the ensemble of trajectories as 
  a function of time. 
  %MRS5: I would put this next section up until just before Eq. 15 in the supporting information (keep any citations in the main text). 
  To derive an analytical equation, first consider an initial pulse spreading out 
  over time with a fixed mean. We can solve for the time-dependent probability density
  of particle positions by solving the one dimensional diffusion equation:
  \begin{equation}
  \frac{\partial p}{\partial t} = D \frac{\partial^2 p}{\partial z^2}
  \end{equation}
  The appropriate initial and boundary conditions are:
  $$BC1: t > 0, z = \infty, p = 0$$
  $$BC2: t > 0, z = 0, \frac{\partial p}{\partial z} = 0$$
  $$IC: t = 0, c = M\delta(z)$$
  where $M$ is the total number of particles in the system. It has been shown 
  elsewhere that the solution to this equation is:~\cite{cussler_diffusion:_2009}
  \begin{equation}
  p(z, t) = \frac{1}{\sqrt{4 \pi D t}}\exp\bigg(\frac{-z^2}{4Dt}\bigg)
  \end{equation}
  We can make the substitution $z = z - vt$, where $v$ represents a constant average
  velocity, in order to linearly shift the mean as a function of time:
  \begin{equation}
  p(z, t) = \frac{1}{\sqrt{4 \pi D t}}\exp\bigg(\frac{-(z - vt)^2}{4Dt}\bigg)
  \end{equation}
  One can track the fraction of particles, $F$, that have crossed the pore 
  boundary by integrating:
  \begin{equation}
  F(t) = \int_L^\infty p~dz = \mathrm{erfc}\bigg(\frac{L - vt}{2\sqrt{D t}}\bigg)
  \end{equation}
  where $L$ is the pore length. This represents the cumulative first passage 
  time distribution so we take its derivative in order to arrive at the first
  passage time distribution:
  \begin{equation}
  P(t) = -\frac{1}{\sqrt{\pi}}e^{-(L - vt)^2 / (4Dt)}\bigg(-\frac{D(L - vt)}{4(Dt)^{3/2}} - \frac{v}{2\sqrt{Dt}}\bigg)
  \label{eqn:passage_times}
  \end{equation} 
  where the only free parameters for fitting are $v$ and $D$. We calculated the
  expected value of Equation~\ref{eqn:passage_times} in order to get the MFPT.

@@

  %MRS5: One question is if other method would lead to substantially
  %different results for fluxes or selectivity.  May need too think
  %about justification more if just one method is mentioned.

  %MRS5: I think it might be useful to have at least visualizions of
  %distribution of FPT and its fit.  It might not be obvious to
  %readers that the different mechanisms have distributions of FPT
  %that are well decscribed by the relatively simple moving diffusion
  %equation in eq 15.

  %Does't fitting to eq 15 cut down on the expense of the number of trajectories
  %by a fair amount?  Should there be some discussion of this?

  %BJC4: not sure if the next couple paragraphs belong in methods 
  %MRS5: this could potentially be in methods.  Alternatives can be discussed there.
  It is computationally infeasible to simulate trajectories long enough that they
  traverse the length of a macroscopic pore. To date, the thinnest H\textsubscript{II}
  LLC membrane synthesized with the monomer in this work was 7$\mu$m thick. Using
  24 cores to simulate trajectory realizations in parallel, it takes on the order 
  of 1 day to simulate 50000 realizations of solutes traversing a 50 nm pore. The 
  RAM requirements and performance scales greater than linearly and thus would 
  take an infeasible amount of memory and time to simulate transport through a 
  pore over 100 times longer.

  %BJC4: methods?  
  %MRS5: This requires knowing the data, so maybe here. 
  We used simulated trajectories which traverse computationally-reasonable length
  pores in order to construct an empirical model which one can use to estimate 
  particle flux for arbitrary length pores. As shown in Figure~\ref{fig:flux_curves},
  the flux appears to scale according to a power law of the form:
  \begin{equation}
  J(L) = cL^{-\beta} 
  \label{eqn:flux_decay}
  \end{equation}
  %while the MFPT scales as its inverse, $(1/c)L^{\beta}$.

  %MRS: can you put a physical thesis here, not just a mathematical one?  Why is this change in parameters important to flux?
  The prefactor, $c$, in Equation~\ref{eqn:flux_decay} grows with increasing
  solute MSD while the value of the growth exponent, $\beta$, is inversely 
  related to the Hurst parameter, $H$ (see Figure~\ref{fig:flux_parameters}).
  In Figure~\ref{fig:flux_curves_brownian}, we removed the correlation 
  structure (set $H$=0.5) from AD trajectory realizations so that all hops 
  were uncorrelated. Regardless of the parameters of the dwell time distribution,
  the growth exponent is always about 2, the same value that we observe for 
  Brownian motion (see Figure~\ref{S-fig:mfpt_curve_brownian} of the 
  Supporting Information).

  %MRS5: probably the figure showing beta \sim 2 can be in supporting
  %information, and you just state the result here.

@@

%MRS: can you connect directly to $\beta$ and $c$ here?  Important to
%make connections between macroscopic parameters and the molecular
%reason for them whenever possible.  Can combine with other paragraphs that talk mainly about c and beta.

@@

  %BJC: need simulations with increased length for L=35-50 nm pores of acetic acid and methanol
  %BJC: see of clauset has better way of fitting t^-alpha (alpha might be too big)
  %BJC: not sure mfpt is necessary. Takes up space.
  %BJC: change alpha to beta
  %BJC: make a chart with parameters
  %BJC: should have a paragraph about c -- roughly follows their MSDs.
  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{flux_curves.pdf}
  \caption{Fits to $cL^{-\beta}$}\label{fig:flux_curves_regular}
  \end{subfigure}
  % BJC: could use better color scheme. Maybe outline hurst parameters
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{flux_parameters.pdf}
  \caption{}\label{fig:flux_parameters}
  \end{subfigure}
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{mfpt_curves.pdf}
%  \caption{}\label{fig:mfpt_curves}
%  \end{subfigure}
  %BJC: going out to 3 decimal places seems silly
  %BJC: might get closer to beta=2 if I use different curve fitting technique than fitting line to log-log
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{flux_curves_brownian.pdf}
  \caption{Fixed $H=0.5$}\label{fig:flux_curves_brownian}
  \end{subfigure}
%  \begin{subfigure}{0.475\textwidth}
%  \includegraphics[width=\textwidth]{mfpt_curves_brownian.pdf}
%  \caption{}\label{fig:mfpt_curves_brownian}
%  \end{subfigure}
  %MRS5: the b) pane is a little messy.  If c) is moved to the supporting, maybe can make b into two figures, and 
  % have each just show 1 point? 
  \caption{(a) The single particle solute flux decays with increasing pore length. 
  Solutes with high MSDs have higher fluxes. (b) We fit the single particle solute flux 
  versus pore length, $L$, to a power law function of the form $cL^{-\beta}$ (dashed 
  lines in (a)). Solutes with higher fluxes have higher values of $c$. Solutes with 
  smaller Hurst parameters, $H$, have larger values of $\beta$. (c) If we remove 
  correlation between hops (set $H$=0.5), the flux increases significantly and 
  $\beta \sim 2$ for all solutes independent of dwell distribution
  parameters.}\label{fig:flux_curves}
  \end{figure}

@@

  %MRS5: figure seem to show that the selectivity plateaus more or less, so thin coatings enough. 
  In Figure~\ref{fig:selectivity}, we plot Equation~\ref{eqn:selectivity_ratio} for pore
  lengths ranging from those studied in Figure~\ref{fig:flux_curves} to macroscopic-length
  pores. Greater differences in $\beta$ values leads to selectivities that are stronger 
  functions of pore length. For $(c_i / c_j)=1$, LLC membranes will be more selective
  towards passage of solutes with lower $\beta$ values versus those with high $\beta$ values.
  When $\beta$ values are equal, as is the case for uncorrelated motion, selectivity only
  depends on the pre-factors, $c$. 
  %MRS5? can you remind people of the physical source (in terms of dwell times, hop lengths) of the differences in beta and c? Want to try to stablish the connection between particle dynamic behavior and selectivity.

@@

  Out of the solutes studied, the data suggests that this particular LLC membrane 
  %MRS5: would is a bit strong given the number of limitations.
  %would 
  might
  be most useful for selectively separating ethylene glycol from methanol and 
  acetic acid. 
  %It also shows reasonable selectivity of Urea over acetic acid.  
  Ethylene glycol has the highest $c$ value and the lowest $\beta$ value. The high
  value of $c$ itself greatly enhances its selective passage. Even more importantly,
  the relatively large difference between its $\beta$ value and those of methanol 
  and acetic acid give rise to a strong pore length dependence. The selectivity is
  enhanced as the pore length increases at the cost of lower solute flux, a 
  phenomenon ubiquitous in polymer separations membranes.~\cite{geise_water_2011}

@@

  %MRS5: sentence fragment left here?
    Neglects concentration polarization, membrane defects and fouling.
